Bow shocks, bow waves, and dust waves. IV. Shell shape statistics

### 2019-05-09 papers/Papers1234/sec-quadrics-observations.tex
Finish cool star discussion and make headway on Orion sources

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{figs/mipsgal-Rc-R90-vs-Orion}
  \caption[]{Shape comparison between Orion Nebula arcs and OB star
    bow shocks.  Other details of the plots are as in
    Figs.~\ref{fig:mipsgal-shapes} and~\ref{fig:mipsgal-uncorrelated},
    except that a larger KDE smoothing bandwidth of
    \(0.06 \times 0.045\) is used for the Orion Nebula sample.}
  \label{fig:ll-compare-mipsgal}
\end{figure}

@@

>>>>

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{figs/mipsgal-Rc-R90-vs-Orion-R0-D}
  \caption[]{Shape comparison between Orion Nebula arcs (green) and OB
    star bow shocks (purple).  Other details of the plots are as in
    Figs.~\ref{fig:mipsgal-shapes} and~\ref{fig:mipsgal-uncorrelated},
    except that a larger KDE smoothing bandwidth of
    \(0.06 \times 0.045\) is used for the Orion Nebula sample.  The colors
    of the individual points for the Orion Nebula sample show the
    size-to-distance ratio, \(R_0/D\), on a logarithmic scale from
    \(R_0/D = 0.002\) (dark green) to \(R_0/D = 0.018\) (light
    green).}
  \label{fig:ll-compare-mipsgal}
\end{figure}

@@

Uniquely among the three bow shock datasets, for the Orion Nebula arcs
we know the source of the external flow.  This is not the stellar wind
from the high-mass stars at the center of the cluster, which has
insufficient momentum at the position of the arcs \citep{Bally:2000a},
nor is it due to stars' own motions, since the space velocities of all
these sources are very low, conforming to the velocity dispersion
\(\sigma_{\text{1D}} \approx \SI{2}{km.s^{-1}}\) of the Orion Nebular Cluster
\citep{Dzib:2017a, Kim:2018b}.  Instead it is due to the transonic
champagne flow of photoionized gas away from the core of the nebula
\citep{Zuckerman:1973a, Henney:2005a}, which has its density peak in
the Orion~S region \citep{Weilbacher:2015a}, located roughly \(30''\)
SW of the dominant Trapezium star \thC.  The bow shock sources are
located at much larger distances of \(D = 90''\) to \(700''\) from
\thC, so that the ratio \(R_0/D\) can be used as a proxy for the local
divergence of the champagne flow on the scale of the bow shock (notice
that the symmetry axes of the bow shocks in Fig.~\ref{fig:ll-arcs} are
always roughly parallel to the direction to \thC{}).  This is
indicated by the color of the points in
Figure~\ref{fig:ll-compare-mipsgal}, with lighter colors corresponding
to a larger \(R_0/D\) and hence a more divergent external flow.  It
can be seen that there is some correlation between the bow shock
shapes and \(R_0/D\): the source with the largest \(\Pi\), 109--246, also
has the largest \(R_0/D\), while the four sources with low \(\Pi\) and
\(\Lambda\) all have low \(R_0/D\).  However, the correlation is far from
perfect, with a Pearson correlation coefficient of only \(r = 0.47\)
between \(\Pi\) and \(R_0/D\).

### 2019-05-09 papers/Papers1234/sec-standing-wave.tex
Finish cool star discussion and make headway on Orion sources

An attractive feature of the oscillation hypothesis is that it
naturally explains why we find little correlation between the bow
shock shape and other source parameters (\S~\ref{sec:corr-shape}).
The one significant correlation that we do find is that the alatude
distribution is broader for bow shocks with larger angular sizes.
This might be explained if the relative amplitude of oscillations were
higher for sources with more powerful winds.  However, in that case it
is hard to understand why no such variation is seen in the width of
the planitude distribution.

We now address the difference in shape distribution between the
different classes of source.  Compared with the OB star bow shocks,
the cool star sample from \citet{Cox:2012a} shows a significantly
smaller alatude of \(\Lambda = 1.41 \pm 0.03\) (see
Fig.~\ref{fig:herschel-compare-mipsgal}).\footnote{Although the
  planitude also appears to be slightly smaller, this is not very
  statistically significant due to the small number of cool star
  sources and the large width of the OB star planitude distribution.}
Such a closed shape for the wings is inconsistent with the wilkinoid
value of \(\Lambda = \text{\numrange{1.63}{1.73}}\), which is surprising
given that the emission shells in these sources are relatively narrow
(Figs.~\ref{fig:herschel-arc-fits}
and~\ref{fig:herschel-arc-fits-poor}), so one might have thought that
the thin-shell approximation of \citet{Wilkin:1996a} would be
\emph{more} appropriate than for the OB stars, but this is clearly not
the case.  One possible explanation for this

>>>>

An attractive feature of the oscillation hypothesis is that it
naturally explains why we find little correlation between the bow
shock shape and other source parameters (\S~\ref{sec:corr-shape}),
since the instantaneous shape at any instant is largely a matter of
chance rather than being due to any intrinsic property of the source.
The one significant correlation that we do find is that the alatude
distribution is broader for bow shocks with larger angular sizes.
This might be explained if the relative amplitude of oscillations were
higher for sources with more powerful winds. \citet{Meyer:2016a} in
their Fig.~2 plot the ``axis ratio'' (which is our \(\Lambda^{-1}\)) as a
function of apex distance for a large set of hydrodynamic bow shock
simulations.  They find the most unstable bow shocks (with the largest
spread in \(\Lambda\)) to be those associated with the highest mass stars in
a relatively dense medium, which have apex distances of
\SIrange{0.3}{1}{pc}.  This corresponds to \(R_0 > 15''\) for
distances less than \SI{4}{kpc}, which is larger than the median
angular size for the MIPSGAL sources, making this a plausible explanation for our statistical result.
% However, in that case it
% is hard to understand why no such variation is seen in the width of
% the planitude distribution.

We now address the difference in shape distribution between the
different classes of source.  Compared with the OB star bow shocks,
the cool star sample from \citet{Cox:2012a}\footnote{Three of these
  sources are high-mass red supergiant (RSG) stars, while the
  remaining 13 are intermediate-mass asymptotic giant branch (AGB)
  stars.} shows a significantly smaller alatude of
\(\Lambda = 1.41 \pm 0.03\) (see
Fig.~\ref{fig:herschel-compare-mipsgal}).\footnote{Although the
  planitude also appears to be slightly smaller, this is not very
  statistically significant due to the small number of cool star
  sources and the large width of the OB star planitude distribution.}
Such a closed shape for the wings is inconsistent with the wilkinoid
value of \(\Lambda = \text{\numrange{1.63}{1.73}}\), which is surprising
given that the emission shells in these sources are relatively narrow
(Figs.~\ref{fig:herschel-arc-fits}
and~\ref{fig:herschel-arc-fits-poor}), so one might have thought that
the thin-shell approximation of \citet{Wilkin:1996a} would be
\emph{more} appropriate than for the OB stars, but this is clearly not
the case.  One possible explanation for this might be that the bow
shocks have not had time to reach a steady-state configuration, as was
suggested by \citet{Mohamed:2012a} for the case of \(\alpha\)~Ori.  The
dynamical timescales, \(R_0 / V\wind\), for the cool star bow shocks
are of order \SI{e4}{yr} and numerical simulations \citetext{e.g.,
  Fig.~11 of \citealp{Mohamed:2012a} and Fig.~2 of
  \citealp{van-Marle:2014a}} show that the bow shock wings take of
order \SIrange{3e4}{e5}{yr} to fully unfold.  However, this is still
short compared with typical RSG and AGB lifetimes, so while it might
apply to a single source it does not work as an explanation for an
entire class of sources, especially given that the alatude and
planitude distributions for these sources are so narrow
(Fig.~\ref{fig:herschel-compare-mipsgal}).  A more promising
explanation is that the shape difference reflects a different origin
for the infrared-emitting dust.  In hot stars, the stellar wind is
dust-free, so the only dust shell comes from the interstellar medium
and lies outside the astropause (contact discontinuity).  In cool
stars, the stellar wind is dusty, which could give rise to significant
infrared emission from near the stellar wind's termination shock.
This is certainly the case for at least one of our cool star sources,
the C-AGB star CW~Leo (IRC+10216), where ultraviolet GALEX
observations \citep{Sahai:2010a} clearly show \emph{both} the outer
shock and the wind termination shock, and comparison with the Herschel
images demonstrate that the infrared dust emission is associated with
the latter.

% However, in order for the dust emission arc to be less open than the
% wilkinoid case, one would require the shocked stellar wind shell to be
% relatively thick, which means that either the Mach number must be low
% or the post-shock cooling must be inefficient (see Paper~I).  In some
% red supergiants, the stellar wind is heated and photoionized by the
% ambient radiation field before reaching the bow shock
% \citep{Morris:1983a, Mackey:2014a}, which would tend to produce a low
% Mach number shock.

In the case of the Orion Nebula bow shocks
(\S~\ref{sec:stat-emiss-line}), we have the opposite situation, where
the alatude and planitude are both significantly larger on average
than for the OB star sources (Fig.~\ref{fig:ll-compare-mipsgal}).

None of these sources have high velocities, but instead follow the
